---
title: "Reproducible Research: Peer Assessment 1"
output: 
    html_document:
        keep_md: true
---

First I load all the packages neccesary.
```{r load_packages, echo = TRUE, results = 'hide'}
library(lubridate)
library(dplyr)

```


## Loading and preprocessing the data



This is the code to load the data and convert date to DATE.  

```{r load_data}
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
activity$date <- ymd(activity$date)

```
  
## What is mean total number of steps taken per day?
First I  collapse the data using the *aggregate* function to find the
total number of steps per day.

```{r collapse}
stepsday <- aggregate(activity$steps, by=(list(activity$date)), FUN=sum, na.rm=TRUE)
```

Then I create a histogram of the steps per day. And report the mean and median.
```{r historgram}
hist(stepsday$x, breaks = 10, col = "darkgreen", main = "Histogram of Steps per Day")

mean(stepsday$x)
median(stepsday$x)
```

**The mean is `r mean(stepsday$x)`. The median is `r median(stepsday$x)`.**  
  
  

## What is the average daily activity pattern?

First it's neccesary collapse the data using the *aggregate* function to find 
the average number of steps per time interval. Then I plot the data.

```{r collapse2}
stepsinterval <- aggregate(activity$steps, by=(list(activity$interval)),
                           FUN=mean, na.rm=TRUE)

with(stepsinterval, plot(Group.1, x, type="l", xlab = "Time interval", 
                         ylab = "Average steps", main = "Average Daily Pattern"))

```

To calculate the maximun I use the *which.max* function.
```{r maxtime}
stepsinterval$Group.1[which.max(stepsinterval$x)]

```


## Imputing missing values
I first use the summary function to find which variables have missing values.

```{r}
summary(activity)
```

Only steps has missing values (2304), so I can count only those.

```{r}
sum(is.na(activity$steps))
```

There are `r sum(is.na(activity$steps)) ` missing values.  

To impute the missing values, I will replace them by the mean of that interval
over all the days.  

```{r}

activity2 <- activity
for(i in 1:length(activity$steps)) {
    if(is.na(activity[i, 1])) {
        index <- i %% 288
        if(index == 0) {index <- 288}
        activity2[i,1] <- stepsinterval[index,2]
    } 
}

stepsday2 <- aggregate(activity2$steps, by=(list(activity2$date)), FUN=sum, na.rm=TRUE)


hist(stepsday2$x, breaks = 10, col = "darkgreen", main = "Histogram of Steps per Day")

mean(stepsday2$x)
median(stepsday2$x)

```

**Both the mean and the median change and become `r mean(stepsday2$x)`.**  


## Are there differences in activity patterns between weekdays and weekends?

I first create a factor variable to see if the day is a weekday or a weekend.
Then I plot the average number of steps per interval using a panel plot.

```{r weekdays_weekends}
typeday <- vector()

for(i in 1:length(activity2$steps)) {
    day <- weekdays(activity2$date[i])
    if(day == "Saturday" || day == "Sunday") {
       typeday[i] <- "weekend"
    } else {
        typeday[i] <- "weekday"
    }
}

typeday <- as.factor(typeday)
activity2 <- cbind(activity2, typeday)


stepsinterval2 <- aggregate(activity2$steps, by=(list(activity2$interval, activity2$typeday)),
                           FUN=mean, na.rm=TRUE)


par(mfrow = c(2,1), mar=c(4,4,1,2))

with(filter(stepsinterval2, Group.2=="weekday"), plot(Group.1, x, type="l", main = "Weekday", xlab = "", ylab = "Average number of steps"))
with(filter(stepsinterval2, Group.2=="weekend"), plot(Group.1, x, type="l", main = "Weekend", xlab = "Interval", ylab = "Average number of steps"))

```

We can see that there are significant differences between weekdays and weekends.
In general, the individual walks more on weekends.


